---
output: github_document
bibliography: vignettes/References.bib
---

<!-- README.md is generated from README.Rmd. Please edit README.Rmd -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```


# GDPTS

<!-- badges: start -->
[![R-CMD-check](https://github.com/MatthewHeun/GDPTS/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/MatthewHeun/GDPTS/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/MatthewHeun/GDPTS/graph/badge.svg)](https://app.codecov.io/gh/MatthewHeun/GDPTS)
<!-- badges: end -->

The goal of GDPTS is to develop a harmonized time series 
of country and world GDP,
starting from the Penn World Tables.
See https://www.rug.nl/ggdc/productivity/pwt/
and
[Feenstra et al. [-@Feenstra:2015aa]](https://doi.org/10.1257/aer.20130954).


## Statement of need

There are many occasions when harmonized 
GDP time series are helpful for analyses.
This package builds that time series
and makes it available via an R package.
In addition, a population time series
is available in this package.


## Installation

You can install the development version of GDPTS from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("MatthewHeun/GDPTS")
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(GDPTS)
## basic example code
```


## More Information

Find more information, including vignettes and function documentation, at 
<https://MatthewHeun.github.io/GDPTS/>.


## References
