---
title: "GDPTS"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{GDPTS}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(GDPTS)
```

```{r}
head(pwt_harmonized)
```




This vignette discusses the harmonization of
population and GDP data from
different versions of the 
Penn World Tables (PWT)
[Feenstra et al. [-@Feenstra:2015aa]](https://doi.org/10.1257/aer.20130954).


## The need for PWT harmonization}

Population and GDP time series from the PWT
for population-adjusted and GDP-adjusted rebound calculations.
(See \ref{sec:re_expressions}.)
However, the current version of the PWT (\pwtnow{}) does not
include population or GDP data for
the Union of Soviet Socialist Republics (USSR),
Yugoslavia (YUG), and Czechoslovakia (CSK)
prior to their dissolutions around 1990.
Thus, a large portion of global population and GDP
for years prior to 1990 are not included
in \pwtnow{}.
Without population and GDP data for years prior to 1990,
a large jump in \%/year variations can be expected
between 1989 and 1990.
Jumps in \%/year time series of population and GDP
can lead to discontinuities in calculated rebound time series.


%++++++++++++++++++++++++++++++
\subsection{Harmonizing PWT data}
\label{sec:harmonizing_pwt_data}
%++++++++++++++++++++++++++++++

Earlier versions of the PWT do, in fact,
include population and GDP time series
for the USSR, Yugoslavia, and Czechoslovakia.
The last PWT version to include those data was \pwtold{},
which used 1985 as the base year~\citep{Summers:1991aa}
for GDP time series.
The most recent version (\pwtnow{})
uses 2017 as the base year~\citep{Feenstra:2015}
for GDP time series.

\pwtold{} and \pwtnow{} provide different real GDP time series
that are comparable
across both countries and years within their respective versions.
We select \texttt{rgdpch} from \pwtold{},
a chained purchase power parity (PPP) time series in 1985\$,
described as ``the recommended intertemporal GDP time series''
in \pwtold{} documentation~\citep[p.~344]{Summers:1991aa}.
We select \texttt{rgdpe} from \pwtnow{},
an expenditure-sided real GDP time series
at chained purchase power parity (PPP) in 2017\$,
described as
``real GDP using prices that are constant across countries and
are also \emph{constant over time}''
(emphasis in original)~\citep[p.~3153]{Feenstra:2015}.
But the \texttt{rgdpch} time series from \pwtold{}
must be converted from base year 1985 to base year 2017
to be consistent with \pwtnow{}.


%++++++++++++++++++++++++++++++
\subsection{Converting \pwtold{} GDP time series from 1985\$ to 2017\$}
\label{sec:pwt56_conversion_to_2017}
%++++++++++++++++++++++++++++++

To convert the \texttt{rgdpch} time series from the 1985 base year of \pwtold{}
to the 2017 base year of \pwtnow{},
we use USA data from both versions.
\texttt{rgdpch} (\pwtold{}) is provided in per-capita terms,
so we first multiply the USA \texttt{rgdpch} time series
by the USA population provided in \pwtold{}.
(See Figure~\ref{fig:PWTPopulationComparisonGraph}
for a comparison of USA population available in \pwtold{} and \pwtnow{}.)

<<PWTPopulationComparisonGraph, eval=eval_R_chunks, fig.cap="Comparison of USA population time series from \\pwtold{} and \\pwtnow{}.">>=
pwt_comparison_df |>
  dplyr::filter(varname %in% c("pop_56", "pop_1001"),
                !is.na(value)) |>
  dplyr::mutate(
    value = value / 1e6,
    varname = dplyr::case_match(
      varname,
      "pop_56" ~ "PWT5.6",
      "pop_1001" ~ "PWT10.01"
    )
  ) |>
  ggplot2::ggplot(mapping = ggplot2::aes(x = year, y = value, colour = varname)) +
  ggplot2::geom_line() +
  ggplot2::scale_y_continuous(name = "USA population [millions]") +
  ggplot2::scale_colour_viridis_d(direction = -1, end = 0.8) +
  ggplot2::labs(x = NULL, colour = NULL) +
  MKHthemes::xy_theme()
@

Next, we calculate the ratio of \texttt{rgdpe} in \pwtnow{}
to \texttt{rgdpch} in \pwtold{},
both in 1985,
the base year for \pwtold{}.
The ratio
(\Sexpr{fancy_print(rgdpch_fix_ratio_for_2017_base_year, fmt = "%.3f")})
gives the conversion factor from
1985\$ (\texttt{rgdpch} in \pwtold{}) to
2017\$ (\texttt{rgdpe} in \pwtnow{}).
Because both time series
(\texttt{rgdpch} in \pwtold{}
and \texttt{rgdpe} in \pwtnow{})
provide real GDP at constant PPP across all countries and years,
the same conversion factor applies to any country's
\pwtold{} \texttt{rgdpch} data in all years.
Figure~\ref{fig:PWTGDPComparisonGraph}
shows USA
\texttt{rgdpe} from \pwtnow{} (in 2017\$),
USA \texttt{rgdpch} from \pwtold{} (in 1985\$), and
a version of USA \texttt{rgdpch} in 2017\$.
Close agreement between \texttt{rgdpch} in 2017\$
and \texttt{rgdpe} in 2017\$ indicates
the appropriateness of our time series choices.

<<PWTGDPComparisonGraph, eval=eval_R_chunks, fig.cap="Comparison of USA GDP time series from \\pwtold{} and \\pwtnow{}.">>=
# GDP graph
pwt_comparison_df |>
  dplyr::filter(varname %in% c("rgdpch_56", "rgdpe_1001", "rgdpch_56_2017"),
                !is.na(value)) |>
  dplyr::mutate(
    value = value / 1e12, # Show GDP in trillions
    varname = dplyr::case_match(
      varname,
      "rgdpch_56" ~ "rgdpch (PWT5.6) [1985$]",
      "rgdpch_56_2017" ~ "rgdpch (PWT5.6) [2017$]",
      "rgdpe_1001" ~ "rgdpe (PWT10.01) [2017$]"
    ),
    varname = factor(varname, levels = c("rgdpe (PWT10.01) [2017$]",
                                         "rgdpch (PWT5.6) [2017$]",
                                         "rgdpch (PWT5.6) [1985$]"))
  ) |>
  ggplot2::ggplot(mapping = ggplot2::aes(x = year,
                                         y = value,
                                         colour = varname,
                                         linetype = varname)) +
  ggplot2::geom_line() +
  ggplot2::scale_y_continuous(name = "USA GDP [trillion$]",
                              limits = c(0, NA)) +
  ggplot2::scale_colour_viridis_d(direction = -1, end = 0.8) +
  ggplot2::scale_linetype_manual(name = NULL,
                                 values = c("solid", "solid", "dashed")) +
  ggplot2::labs(x = NULL, colour = NULL) +
  MKHthemes::xy_theme()
@


After converting the \pwtold{} \texttt{rgdpch} time series for the USSR,
Yugoslavia, and Czechoslovakia to 2017\$ in the years 1971--1989,
we sum all countries at each year to
obtain world population and GDP estimates,
as shown in Figures~\ref{fig:PWTWorldPopulationGraph}
and~\ref{fig:PWTWorldGDPGraph}, respectively.
The data represented by solid lines
in Figures~\ref{fig:PWTWorldPopulationGraph}
and~\ref{fig:PWTWorldGDPGraph}
are used throughout this paper for calculating
energy and exergy per person,
energy and exergy per GDP, and
rates of change in \%/year of both.
(See Figures~\ref{fig:PercChangeGraph}
and~\ref{fig:PercChangeGraphCountries}.)


<<PWTWorldPopulationGraph, eval=eval_R_chunks, fig.width=6, fig.cap="Harmonized world population and population from \\pwtnow{}~\\citep{Summers:1991aa, Feenstra:2015}.">>=
pwt1001_pop <- pwt10::pwt10.01 |>
  dplyr::select(year, isocode, pop) |>
  dplyr::rename(Year = year) |>
  dplyr::mutate(
    pop = ifelse(is.na(pop), 0, pop),
    pop_1001 = pop * 1e6, # pop is in millions, pop1001 is in persons
    pop = NULL
  ) |>
  dplyr::group_by(Year) |>
  dplyr::summarise(pop_1001 = sum(pop_1001))
# Look at the results
pwt_harmonized |>
  dplyr::filter(Country == "World") |>
  dplyr::select(Year, pop) |>
  dplyr::rename(pop_harmonized = pop) |>
  dplyr::left_join(pwt1001_pop, by = "Year") |>
  tidyr::pivot_longer(cols = dplyr::all_of(c("pop_harmonized", "pop_1001")), names_to = "pop_type", values_to = "pop") |>
  dplyr::filter(Year >= 1971) |>
  dplyr::mutate(
    pop = pop / 1e9, # convert from persons to billion
    pop_type = dplyr::case_match(
      pop_type,
      "pop_1001" ~ "PWT 10.01",
      "pop_harmonized" ~ "Harmonized"),
    pop_type = factor(pop_type, levels = c("Harmonized", "PWT 10.01"))
  ) |>
  ggplot2::ggplot(mapping = ggplot2::aes(x = Year, y = pop, linetype = pop_type)) +
  ggplot2::geom_line() +
  ggplot2::scale_x_continuous(name = NULL) +
  ggplot2::scale_y_continuous(name = "World population [billion]",
                              limits = c(0, 8)) +
  ggplot2::scale_linetype_manual(name = NULL,
                                 values = c("solid", "dashed")) +
  MKHthemes::xy_theme()
@


<<PWTWorldGDPGraph, eval=eval_R_chunks, fig.width=6, fig.cap="Harmonized world GDP and world GDP from \\pwtnow{}~\\citep{Summers:1991aa, Feenstra:2015}.">>=
pwt1001_rgdpe <- pwt10::pwt10.01 |>
  dplyr::select(year, isocode, rgdpe) |>
  dplyr::rename(Year = year) |>
  dplyr::mutate(
    rgdpe = ifelse(is.na(rgdpe), 0, rgdpe),
    rgdpe_1001 = rgdpe * 1e6, # rgdpe in million 2017$, rgdpe_1001 in $
    rgdpe = NULL
  ) |>
  dplyr::group_by(Year) |>
  dplyr::summarise(rgdpe_1001 = sum(rgdpe_1001))

pwt_harmonized |>
  dplyr::filter(Country == "World", Year >= 1971) |>
  dplyr::select(Year, rgdpe) |>
  dplyr::left_join(pwt1001_rgdpe, by = "Year") |>
  tidyr::pivot_longer(cols = dplyr::all_of(c("rgdpe", "rgdpe_1001")), names_to = "rgdpe_type", values_to = "rgdpe") |>
  dplyr::mutate(
    rgdpe = rgdpe/1e12, # convert from 2017$ to trillion2017$
    rgdpe_type = dplyr::case_match(
      rgdpe_type,
      "rgdpe" ~ "Harmonized",
      "rgdpe_1001" ~ "PWT 10.01"),
    rgdpe_type = factor(rgdpe_type, levels = c("Harmonized", "PWT 10.01"))
  ) |>
  ggplot2::ggplot(mapping = ggplot2::aes(x = Year, y = rgdpe, linetype = rgdpe_type)) +
  ggplot2::geom_line() +
  ggplot2::scale_x_continuous(name = NULL) +
  ggplot2::scale_y_continuous(name = "World GDP [trillion 2017$]",
                              limits = c(0, NA)) +
  ggplot2::scale_linetype_manual(name = NULL,
                                 values = c("solid", "dashed")) +
  MKHthemes::xy_theme()
@



## References
